Shiny App Manuscript Figures: Aquatic Organisms

### 2021-05-20 Shiny_Manuscript_Figures/Manuscript_Figures.Rmd
Minor data set changes (just character to numerical values to make plotting easier) + implementation of screening plot + random scripts for possible figure generation for paper

---
title: 'Shiny App Manuscript Figures: Aquatic Organisms'
author: "Leah Thornton Hampton"
date: "4/2/2021"
output:
  html_document: 
    toc: true
    toc_depth: 4
    number_sections: true
    toc_float: true
---

@@

#Import packages
library(tidyverse)

@@

#Data Tidying
aoc_v1 <- aoc %>% # start with original dataset
   # full dataset filters.
  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
    effect == "N" ~ "No"),
    levels = c("No", "Yes"))) %>%
  # removing NAs to make data set nicer
  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported", chem.exp.typ.nominal = "Particle Only"))

@@

    size.category == 0 ~ "Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.

@@

                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed")))

@@

```{r Organism Groups, echo = FALSE, results = 'hide', error = FALSE, warning = FALSE, message = FALSE}

a <- aoc_setup %>% 
  filter(environment != "Terrestrial") %>% 
  ggplot(aes(x = org_f, fill = environment)) +
  geom_bar()

plot(a)
```

>>>>

---
title: 'Team X Manuscript Figures: Aquatic Organisms'
author: "Leah Thornton Hampton"
date: "4/26/2021"
output:
  html_document: 
    toc: true
    toc_depth: 4
    number_sections: true
    toc_float: true
---

@@

#Import packages
library(tidyverse)
library(collapsibleTree) #plot type for endpoint category tree
library(calecopal) #colors 
library(wesanderson) #colors

@@

#Data Tidying
# Master dataset for scatterplots - for Heili's tab.
aoc_v1 <- aoc %>% # start with original dataset
   # full dataset filters.
  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
    effect == "N" ~ "No"),
    levels = c("No", "Yes"))) %>%
  # removing NAs to make data set nicer
  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported", chem.exp.typ.nominal = "Particle Only"))

@@

    size.category == 0 ~ "Size Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Size Not Reported"))) %>% # creates new column with nicer names and order by size levels.

@@

                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed"))) %>% 
  #Remove leachate and additive/chemical transfer experiments
  dplyr::filter(leachate.only != "Y") %>%
  mutate(chem.exp.typ.nominal_f = factor(case_when(chem.exp.typ.nominal == "Particle Only" ~ "Particle Only",
                                                   chem.exp.typ.nominal == "co.exp" ~ "Chemical Co-Exposure",
                                                   chem.exp.typ.nominal == "sorbed" ~ "Chemical Transfer"))) %>% 
  dplyr::filter(chem.exp.typ.nominal_f == "Particle Only") %>% 
  #calculate maximum ingestible size (if not already in database)
  mutate(max.size.ingest.mm = ifelse(is.na(max.size.ingest.mm), 
    10^(0.9341 * log10(body.length.cm - 1.1200)) * 10,  #(Jamm et al 2020 Nature paper)correction for cm to mm
    max.size.ingest.mm)) # if already present, just use that

@@

```{r collapsible tree example}

collapsibleTree(aoc_setup, root = "Aquatic Organisms Database", hierarchy = c("lvl1_f", "lvl2_f", "lvl3_f", "bio_f"),
                    fontSize = 16, zoomable = FALSE,    
    fill = c(
      # The root
      "seashell",
      # lvl1
      rep("turquoise", length(unique(aoc_setup$lvl1_f))),
      # lvl2
      rep("palegreen", length(unique(paste(aoc_setup$lvl1_f, aoc_setup$lvl2_f)))),
      # lvl3
      rep("hotpink", length(unique(paste(aoc_setup$lvl1_f, aoc_setup$lvl2_f, aoc_setup$lvl3_f)))),
      # bio org
      rep("orchid", length(unique(paste(aoc_setup$lvl1_f, aoc_setup$lvl2_f, aoc_setup$lvl3_f, aoc_setup$bio_f))))))

```{r endpoints and organism groups, fig.width = 15}
library(viridis)
a <- aoc_setup %>% 
  filter(effect_f != "NA") %>%
  count(lvl1_f, org_f) %>% #how many obs of each organism group within each lvl1
  group_by(lvl1_f) %>%
  mutate(n_lvl1 = sum(n)) %>% #how many obs within each lvl1
  ungroup() %>%   
  ggplot(aes(x = fct_reorder(lvl1_f, n_lvl1), y = n, group = n, fill = org_f)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), color = "black", width = 1) +
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 10)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values= wes_palette("Zissou1", type = "continuous", 13)) +
  theme_classic() +
  theme(text = element_text(size = 16))+
  labs(y = "Number of Endpoints in Database", x = "Endpoint Category", fill = "Organism\nGroup")

plot(a)

```{r particle characteristics}

particle <- aoc_setup %>%
  group_by(size_f) %>%
  #count the number of data points in each size category
  mutate(size_count = n()) %>% 
  #percentage that report polymer
  mutate(poly_pct = (sum(poly_f != "Not Reported")/size_count)*100) %>% 
  #percentage that report shape
  mutate(shape_pct = (sum(shape_f != "Not Reported")/size_count)*100) %>%
  #percentage that report density
  replace_na(list(density.reported.estimated = "Not Reported")) %>%   
  mutate(density_f = as.factor(density.reported.estimated)) %>% 
  mutate(density_pct = (sum(density_f == "reported")/size_count)*100) %>%  
  #percentage that report charge
  mutate(charge_pct = (sum(!is.na(charge))/size_count)*100) %>%
  #percentage that report functional group
  mutate(fxngroup_pct = (sum(!is.na(functional.group))/size_count)*100) %>%
  #validate size
  mutate(sizevalid_f = as.factor(size.valid)) %>% 
  mutate(sizevalid_pct = (sum(sizevalid_f == "Y")/size_count)*100) %>% 
  #validate shape
  mutate(shapevalid_f = as.factor(shape.valid)) %>% 
  mutate(shapevalid_pct = (sum(shapevalid_f == "Y")/size_count)*100) %>% 
  #validate polymer
  mutate(polyvalid_f = as.factor(polymer.valid)) %>% 
  mutate(polyvalid_pct = (sum(polyvalid_f == "Y")/size_count)*100) %>% 
  ungroup() %>% 
  group_by(size_f, poly_pct, shape_pct, density_pct, charge_pct, fxngroup_pct, sizevalid_pct, shapevalid_pct, polyvalid_pct) %>% 
  summarise() %>% 
  rename("Polymer Type" = poly_pct, "Morphology" = shape_pct, "Density" = density_pct, "Charge" = charge_pct, "Functional Group" = fxngroup_pct, "Size Validated" = sizevalid_pct, "Morphology Validated" = shapevalid_pct, "Polymer Validated" = polyvalid_pct)

require(pheatmap)
require(RColorBrewer)

# Make column 1 equal to the row name and remove data from column 1 and arrange size levels

All_data <- data.frame(particle, row.names = 1)%>% 
  arrange("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Size Not Reported")

#Turn data into a matrix format

All_matrix <- data.matrix(All_data)

color <- brewer.pal(n=8, name = "Blues")

Pretty_heatmap <- pheatmap(All_matrix,
        main = "Frequency (%) of Particle Characteristics Reported",
        fontsize = 12,
        treeheight_row = 0, treeheight_col = 0,
        cluster_rows = FALSE, cluster_cols = FALSE,
        display_numbers = TRUE,
        color = color,
        number_format = "%.0f", fontsize_number = 10, number_color = "black")

### 2021-05-20 TeamX_Manuscript_Figures/Manuscript_Figures_TeamX.Rmd
Minor data set changes (just character to numerical values to make plotting easier) + implementation of screening plot + random scripts for possible figure generation for paper

---
title: 'Team X Manuscript Figures: Aquatic Organisms'
author: "Leah Thornton Hampton"
date: "4/26/2021"
output:
  html_document: 
    toc: true
    toc_depth: 4
    number_sections: true
    toc_float: true
---

@@

```{r Data Import and Packages, echo = FALSE, results = 'hide', error = FALSE, warning = FALSE, message = FALSE}

#Import packages
library(tidyverse)
library(collapsibleTree) #plot type for endpoint category tree
library(calecopal) #colors 
library(wesanderson) #colors

@@

    size.category == 0 ~ "Size Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Size Not Reported"))) %>% # creates new column with nicer names and order by size levels.

@@

```{r collapsible tree example}

collapsibleTree(aoc_setup, root = "Aquatic Organisms Database", hierarchy = c("lvl1_f", "lvl2_f", "lvl3_f", "bio_f"),
                    fontSize = 16, zoomable = FALSE,    
    fill = c(
      # The root
      "seashell",
      # lvl1
      rep("turquoise", length(unique(aoc_setup$lvl1_f))),
      # lvl2
      rep("palegreen", length(unique(paste(aoc_setup$lvl1_f, aoc_setup$lvl2_f)))),
      # lvl3
      rep("hotpink", length(unique(paste(aoc_setup$lvl1_f, aoc_setup$lvl2_f, aoc_setup$lvl3_f)))),
      # bio org
      rep("orchid", length(unique(paste(aoc_setup$lvl1_f, aoc_setup$lvl2_f, aoc_setup$lvl3_f, aoc_setup$bio_f))))))

```{r endpoints and organism groups, fig.width = 15}
library(viridis)
a <- aoc_setup %>% 
  filter(effect_f != "NA") %>%
  count(lvl1_f, org_f) %>% #how many obs of each organism group within each lvl1
  group_by(lvl1_f) %>%
  mutate(n_lvl1 = sum(n)) %>% #how many obs within each lvl1
  ungroup() %>%   
  ggplot(aes(x = fct_reorder(lvl1_f, n_lvl1), y = n, group = n, fill = org_f)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), color = "black", width = 1) +
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 10)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values= wes_palette("Zissou1", type = "continuous", 13)) +
  theme_classic() +
  theme(text = element_text(size = 16))+
  labs(y = "Number of Endpoints in Database", x = "Endpoint Category", fill = "Organism\nGroup")

@@

```{r particle characteristics}

particle <- aoc_setup %>%
  group_by(size_f) %>%
  #count the number of data points in each size category
  mutate(size_count = n()) %>% 
  #percentage that report polymer
  mutate(poly_pct = (sum(poly_f != "Not Reported")/size_count)*100) %>% 
  #percentage that report shape
  mutate(shape_pct = (sum(shape_f != "Not Reported")/size_count)*100) %>%
  #percentage that report density
  replace_na(list(density.reported.estimated = "Not Reported")) %>%   
  mutate(density_f = as.factor(density.reported.estimated)) %>% 
  mutate(density_pct = (sum(density_f == "reported")/size_count)*100) %>%  
  #percentage that report charge
  mutate(charge_pct = (sum(!is.na(charge))/size_count)*100) %>%
  #percentage that report functional group
  mutate(fxngroup_pct = (sum(!is.na(functional.group))/size_count)*100) %>%
  #validate size
  mutate(sizevalid_f = as.factor(size.valid)) %>% 
  mutate(sizevalid_pct = (sum(sizevalid_f == "Y")/size_count)*100) %>% 
  #validate shape
  mutate(shapevalid_f = as.factor(shape.valid)) %>% 
  mutate(shapevalid_pct = (sum(shapevalid_f == "Y")/size_count)*100) %>% 
  #validate polymer
  mutate(polyvalid_f = as.factor(polymer.valid)) %>% 
  mutate(polyvalid_pct = (sum(polyvalid_f == "Y")/size_count)*100) %>% 
  ungroup() %>% 
  group_by(size_f, poly_pct, shape_pct, density_pct, charge_pct, fxngroup_pct, sizevalid_pct, shapevalid_pct, polyvalid_pct) %>% 
  summarise() %>% 
  rename("Polymer Type" = poly_pct, "Morphology" = shape_pct, "Density" = density_pct, "Charge" = charge_pct, "Functional Group" = fxngroup_pct, "Size Validated" = sizevalid_pct, "Morphology Validated" = shapevalid_pct, "Polymer Validated" = polyvalid_pct)

require(pheatmap)
require(RColorBrewer)

# Make column 1 equal to the row name and remove data from column 1 and arrange size levels

All_data <- data.frame(particle, row.names = 1)%>% 
  arrange("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Size Not Reported")

#Turn data into a matrix format

All_matrix <- data.matrix(All_data)

color <- brewer.pal(n=8, name = "Blues")

Pretty_heatmap <- pheatmap(All_matrix,
        main = "Frequency (%) of Particle Characteristics Reported",
        fontsize = 12,
        treeheight_row = 0, treeheight_col = 0,
        cluster_rows = FALSE, cluster_cols = FALSE,
        display_numbers = TRUE,
        color = color,
        number_format = "%.0f", fontsize_number = 10, number_color = "black")

>>>>

---
title: 'Shiny App Manuscript Figures: Aquatic Organisms'
author: "Leah Thornton Hampton"
date: "4/26/2021"
output:
  html_document: 
    toc: true
    toc_depth: 4
    number_sections: true
    toc_float: true
    code_folding: hide
---

@@

```{r Data Import and Packages, warning = FALSE, message = FALSE}

#Import packages
library(tidyverse)

@@

    size.category == 0 ~ "Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.

@@

# Ludo

```{r Ludo, warning = FALSE, message = FALSE, fig.height = 7, fig.width = 12}

a <- aoc_setup %>% 
  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
  filter(shape_f != "Not Reported") %>% #drop studies where shapes aren't reported
  filter(effect_f != "NA") %>% #drop ECX values
  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
  ggplot(aes(y = shape_f, x = dose.mg.L.master, fill = effect_f)) +
  geom_boxplot(alpha = 0.8) +
  scale_x_log10() +
  labs(y = "Morphology", x = "mg/L", fill = "Effect") +
  theme_classic() +
  theme(text = element_text(size = 16))

@@

b <- aoc_setup %>% 
  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
  filter(shape_f != "Not Reported") %>% #drop studies where shapes aren't reported
  filter(effect_f != "NA") %>% #drop ECX values
  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
  ggplot(aes(y = shape_f, x = dose.particles.mL.master, fill = effect_f)) +
  geom_boxplot(alpha = 0.8) +
  scale_x_log10() +
  labs(y = "Morphology", x = "particles/mL", fill = "Effect") +
  theme_classic() +
  theme(text = element_text(size = 16))

plot(b)

c <- aoc_setup %>% 
  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
  filter(!poly_f %in% c("Not Reported", "Polyethylene Vinyl Acetate", "Polycarbonate", "Latex")) %>% 
  filter(effect_f != "NA") %>% #drop ECX values
  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
  ggplot(aes(y = poly_f, x = dose.mg.L.master, fill = effect_f)) +
  geom_boxplot(alpha = 0.8) +
  scale_x_log10() +
  labs(y = "Polymer", x = "mg/L", fill = "Effect") +
  theme_classic() +
  theme(text = element_text(size = 16)) +
  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 10))

plot(c)

d <- aoc_setup %>% 
  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
  filter(!poly_f %in% c("Not Reported", "Polyethylene Vinyl Acetate", "Polycarbonate", "Latex")) %>%
  filter(effect_f != "NA") %>% #drop ECX values
  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
  ggplot(aes(y = poly_f, x = dose.particles.mL.master, fill = effect_f)) +
  geom_boxplot(alpha = 0.8) +
  scale_x_log10() +
  labs(y = "Polymer", x = "particles/mL", fill = "Effect") +
  theme_classic() +
  theme(text = element_text(size = 16)) +
  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 10))

plot(d)
```
